```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

## Identifying Schools with Biggest Differences in Math Test Scores

### Introduction

In this analysis, we will create a scatterplot to visualize the relationship between the FARMS (Free and Reduced-price Meal System) percentage and the percentage of students who met or exceeded math expectations. We'll then use linear regression to identify the schools that show the biggest differences between their actual and expected math test scores.

This analysis can be useful for journalists to understand which schools are performing better or worse than expected, based on the socioeconomic status of their student population. By identifying these schools, targeted interventions and support can be provided to help improve student outcomes - and to provide some context on comments by policymakers.

### Loading the Data

First, let's load the data into a data frame using the read_csv() function from the tidyverse library.

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_pa.csv")
```
### Task: Create a Scatterplot and Describe the Results (2 points)

Next, let's create a scatterplot to visualize the relationship between the FARMS percentage and the percentage of students who met or exceeded math expectations. We'll use ggplot2 to create the plot and add a linear regression line.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations", title = "Scatterplot of Math Expectations vs FARMS Percentage")
```

The scatterplot shows the overall trend of the relationship between the FARMS percentage and the math test scores. The red line represents the linear regression line, which helps us understand the average relationship between the two variables. Based on what you see here, what's the relationship between FARMS percentage and math test scores?

ANSWER HERE

The red line shows a small upward trend, so as FARMS percentage goes up, math scores go up a little. But the dots are spread out, so the connection is weak. That means FARMS percentage doesn’t explain much about the math scores.

### Task: Linear Regression Model (3 points)

Next, we'll perform a linear regression analysis using the lm() function. We'll set the percentage of students who met or exceeded math expectations as the dependent variable, and the FARMS percentage as the independent variable. In other words, if we know the FARMS percentage for a school, how reliably can we predict its math score?

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)

summary(model)
```

There are three things we need to examine here: the p-value, the Adjusted R-squared value and the coefficients. You know what a p-value is and how to evaluate it. The Adjusted R-squared is a measure of how much of the difference between schools' math scores can be explained by their FARMS percentage. You don't need to multiply that figure by anything: that's a percentage. The closer the adjusted R-squared is to 100, the more of the outcome the independent variable explains. If both the p-value and adjusted R-squared values are significant, we can look to the coefficients.

In the middle, you can see the (Intercept) is 59.57897 and the FARMS coefficient is 0.12349. Remember high school algebra? Remember learning the equation of a line? Remember swearing that learning y=mx+b is stupid because you’ll never need it again? Surprise. It’s useful again. In this case, we could try to predict a school's score math score by using y=mx+b. In this case, y is the math score, m is 0.12349 and b is 59.57897. So we would multiply a school's FARMS percentage by 0.12349 and then add 59.57897 to it. The result would tell you what the math score would be, according to our model.

Based on what you've learned so far, is this model worth using? What does it tell us about the nature of the relationship between FARMS percentage and math scores?

ANSWER HERE

Based on the linear regression results, this model is not very strong or reliable. The p-value for the FARMS percentage is 0.123, which is above the cutoff of 0.05. This means the relationship between FARMS percentage and math scores is not statistically significant, we can't be confident that the pattern we see isn’t just chance. 

The adjusted R-squared is only 0.0254, meaning the model explains just about 2.5% of the differences in math scores between schools. While the regression line suggests a small positive trend (as FARMS percentage goes up, math scores slightly increase), the effect is very weak. 

Overall, this model doesn’t do a good job of predicting math scores, and FARMS percentage alone doesn’t explain much of what’s going on.

### Task: Identifying Schools with Biggest Residuals (2 points)

To find the schools with the biggest differences between their actual and expected math test scores, we'll extract the residuals from the linear regression model. The residuals represent the difference between the actual and predicted values. Residuals can tell you several things, but most important is if a linear model the right model for your data. If the residuals appear to be random, then a linear model is appropriate. If they have a pattern, it means something else is going on in your data and a linear model isn’t appropriate.

Residuals can also tell you who is under-performing and over-performing the model. And the more robust the model – the better your r-squared value is – the more meaningful that label of under or over-performing is. 

We'll then use the arrange() function to sort the data frame by the absolute value of the residuals.

```{r}
schools$predicted <- predict(model)

# Calculate the residuals
schools <- schools |> 
  mutate(residual = `% Met or Exceeded Math Expectations` - predicted)

# Sort the data frame by the absolute value of the residuals in descending order
schools <- schools |>
  mutate(abs_residual = abs(residual)) |>
  arrange(desc(abs_residual))

```
The schools with the biggest positive residuals are performing better than expected, while the schools with the biggest negative residuals are performing worse than expected, based on their FARMS percentage. Look at the absolute values of the residuals; what are the largest differences between the predicted result and the actual math scores?

ANSWER HERE

Some schools had big differences between their real math scores and what was predicted based on the number of low-income students (FARMS percentage). Most of these schools did worse than expected. Whitemarsh Elementary had the biggest drop, scoring about 35 points lower than predicted. Other schools like Gulph and Lower Gwynedd Elementary also scored over 30 points below what the model expected. This means these schools may be facing challenges that the FARMS percentage doesn't fully explain.

However, a few schools did better than expected. Crooked Billet and Woodland Elementary scored more than 25 points higher than predicted. This shows that something at these schools is helping students do well in math, even if they have a higher number of low-income families.

## Task: Evaluate the model (3 points)

A school board member has said at a meeting that elementary schools with fewer students on free and reduced price meals do better on standardized math tests. Describe what conclusions you'd draw from this model's results and how it would influence a story pitch about the school board member and math scores.

ANSWER HERE

The school board member said that schools with fewer students getting free or reduced-price meals do better on math tests. The model mostly agrees with that—schools with fewer low-income students often have higher predicted math scores. But the actual results show that this isn’t always true.

Some schools with more low-income students did much better than expected, like Crooked Billet and Woodland Elementary. And some schools with fewer low-income students, like Whitemarsh and Lower Gwynedd, did much worse than predicted. This means other things—like good teaching, strong leadership, or extra support—might be helping students succeed, even if they come from lower-income families.

For a story, this could lead to a pitch like: 
“Is Income the Whole Story? Some Schools Beat the Odds on Math Tests.” 
The story could look at what’s working in the schools that are doing better than expected.
