
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
```
I loaded the necessary packages so I can work with data (read it in, clean it, analyze it, and visualize it)
These are standard R packages that help me handle the data in the way I want.

```{r}
data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQ_02vV4-U327jbHQCTXZGpnDUyW46hKPwMJwVVZXcLeefOo2JR6iR5SPs-GuiK9ZfekcUfvwbTp8eJ/pub?output=csv")
```
I read in the dataset from a public Google Sheet link I created so I can start working with it.
The data loaded into R and is ready to use.

```{r}
data <- data %>%
  mutate(
    Revenue_num = str_remove_all(Revenue, "[$,]") %>% as.numeric()
  )

```
I removeed dollar signs and commas from the revenue column so I can treat the numbers as numeric values for math and analysis. This lets me calculate averages, run tests, and create graphs using revenue data.


```{r}
data_clean <- data %>%
  mutate(
    Year = as_factor(Year),
    Category_Type = case_when(
      str_detect(Category, "Total") ~ "Total",
      str_detect(Category, "Streaming|Subscription|SoundExchange") ~ "Streaming",
      str_detect(Category, "CD|LP|EP|Physical|Video") ~ "Physical",
      TRUE ~ "Other"
    ),
    Category = fct_reorder(Category, Revenue_num, .fun = mean)
  )

```
I cleaned up the dataset by making year a factor and grouping categories into broader types like "Streaming" or "Physical." This helps me simplify the data so it's easier to analyze trends across similar groups.


```{r}
data_analysis <- data_clean %>%
  filter(!str_detect(Category, "Total"))

```
I removed the “Total” rows because I don’t want to count summary values in my analysis. This keeps my results from being skewed by totals that combine other categories.



```{r}
anova1 <- aov(Revenue_num ~ Category, data = data_analysis)
summary(anova1)

```
I ran a one-way ANOVA test to see if revenue is different across categories. If the p-value is low, it means at least one category has different average revenue from the others.




```{r}
anova2 <- aov(Revenue_num ~ Category * Year, data = data_analysis)
summary(anova2)

```
I then ran a two-way ANOVA to check if revenue differences depend on both category and year.This tells me whether trends over time change depending on the category, and if any interactions are significant.



```{r}
library(janitor)

data_summary <- data_analysis %>%
  group_by(Category) %>%
  summarise(
    mean_revenue = mean(Revenue_num, na.rm = TRUE),
    sd_revenue = sd(Revenue_num, na.rm = TRUE),
    .groups = "drop"
  )

```
I calculated the average and standard deviation of revenue for each category to better understand their typical values and spread. This gives context to how high or low each revenue number is for that group.



```{r}
data_analysis <- data_analysis %>%
  left_join(data_summary, by = "Category") %>%
  mutate(z_score = (Revenue_num - mean_revenue) / sd_revenue)

```
I added a z-score column so I can see how each revenue value compares to its category’s average. The Z-scores help me quickly spot unusually high or low revenue values



```{r}
data_corr <- data_analysis %>%
  mutate(Year_num = as.numeric(as.character(Year)))  # convert factor back to numeric

```
I converted the Year variable back into a number so I can use it in a correlation test. I needed numeric values to measure how revenue changes over time.



```{r}
cor_test <- cor.test(data_corr$Revenue_num, data_corr$Year_num)
print(cor_test)

```
I checked if there’s a general trend in revenue over the years across all categories. The correlation result shows whether revenue is rising or falling with time, and if the relationship is strong or weak.



```{r}
library(broom)

category_corrs <- data_corr %>%
  group_by(Category) %>%
  summarise(cor_test = list(cor.test(Revenue_num, Year_num))) %>%
  mutate(
    tidy = map(cor_test, broom::tidy)
  ) %>%
  unnest(tidy)

print(category_corrs)

```
I ran separate correlation tests for each category to see how their revenues change over time. This helps me understand which categories are growing or shrinking and how strongly.




```{r}
ggplot(data_corr, aes(x = Year_num, y = Revenue_num, color = Category_Type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Revenue Trends by Category Type",
    x = "Year",
    y = "Revenue ($)",
    color = "Category Type"
  ) +
  theme_minimal()
```
I lastly made a scatter/line plot showing how revenue changes by year, colored by category type, with trend lines to show patterns. This visual makes it easy to spot which types of music revenue are rising or falling over time.





Main Findings:
The music industry makes the most money from streaming, especially paid subscriptions like Spotify Premium or Apple Music. While CDs and vinyl records are still around, they don’t bring in much money anymore. The data shows streaming is leading the way—and the stats back it up.



What the Data Shows:
1. Paid Streaming Is Driving Growth
Paid streaming has the strongest link to rising music industry revenue.

Correlation: r = 0.994 — this means paid streaming and total revenue go up together almost perfectly.

When more people pay for music streaming, the industry earns more money.


2. Ad-Supported Streaming Also Helps
Free services with ads are also growing.

Correlation: r = 0.983 — strong positive connection, just not quite as strong as paid subscriptions.


3. CDs and Vinyl Aren’t Making a Big Difference
CD revenue has a negative correlation with total industry revenue: r = -0.347.

This means CD sales are slightly going down as total revenue goes up.

Vinyl has a weak positive correlation: r = 0.122, but it’s not strong enough to drive industry growth.



What the ANOVA Test Shows:
ANOVA helps me see if different types of music revenue (like CDs, streaming, vinyl) have real differences in how they affect total revenue.

The test showed that there is a statistically significant difference between the groups (p < 0.001). This shows that some revenue types (like streaming) matter much more than others (like CDs or vinyl).

This confirms what the correlation results showed—streaming is doing most of the heavy lifting, while physical formats don’t add much.



Graph:
The scatter/line graph shows that streaming is rising fast while physical formats stay flat.



Questions:
Why do some people still buy CDs or vinyl, even when streaming is easier and more popular?

How do musicians make money now compared to the days of physical albums?

Are there age or cultural differences in how people listen to music?

What happens to record stores and collectors as streaming takes over?



Reporting the Story:
Double-check the data with sources like IFPI and going deeper into RIAA.



Interview people:

Music business experts to talk about the shift to streaming.

Local artists to share their experience with streaming platforms.

A record store owner or vinyl collector to explain why some fans still love physical formats.

Young and older music fans for a range of perspectives.



Visuals:

Graphs that clearly show how streaming impacts revenue compared to CDs and vinyl. Like the graph I had created here.


Title Ideas:
“Streaming’s Takeover: The Numbers Don’t Lie”

“CDs, Vinyl, and the Rise of Paid Subscriptions”

“Spotify Killed the CD Star: A Deep Dive into Music Revenue”

