---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <-  read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |>mutate(violent_rate_2019 = (violent2019 / pop2019) * 10000,
property_rate_2019 = (property2019 / pop2019) * 10000,violent_rate_2020 = (violent2020 / pop2020) * 10000,property_rate_2020 = (property2020 / pop2020) * 10000)

```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

ANSWER HERE

I noticed that Baltimore City had the highest violent crime in 2019 and Worcester had the highest property crime. In 2020 Worcester had the highest in both violent and property crime. Baltimore had the highest violent crime in 2019, but Worcester took the lead in both violent and property crime in 2020. This shift suggests crime isn’t just an urban issue—smaller areas can see spikes too.

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate(violent_rate_change = ((violent_rate_2020 - violent_rate_2019) / violent_rate_2019) * 100,property_rate_change = ((property_rate_2020 - property_rate_2019) / property_rate_2019) * 100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

ANSWER HERE

Prince George's had the highest violent rate change at 895.32, and Somerset had the highest property rate change at 320.80. Such extreme jumps suggest significant issues rather than gradual trends. Another piece of data to point out is the number of cities that saw a -100 rate change. Could suggest a change in enforcement.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE

Violent crime in Prince George’s County jumped 895.32% from 2019 to 2020, while property crime in Somerset County rose 320.80%, the biggest increases in Maryland. 

These sharp spikes suggest something more than just normal crime trends. Due to things like the pandemic, economic struggles, or changes in law enforcement, this could have played a role with these big jumps. With crime shifting in drastic/unexpected ways, understanding why these jumps happened is key to keeping communities safe.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
