---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
m_dataset <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```



### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10.
```{r}
sample_5 <- m_dataset |> sample_n(5)
sample_10 <- m_dataset |> sample_n(10)
```


### Task 4: Calculate the mean and standard deviation of the samples, using one of the colors.
```{r}
sample_5 |> summarise(mean = mean(red), sd = sd(red))
sample_10 |> summarise(mean = mean(red), sd = sd(red))
```


### Task 5: Calculate the mean and standard deviation of the population, using the same color as Task 4
```{r}
m_dataset |> summarise(mean = mean(red), sd = sd(red))
```




### Task 6: Compare the sample and population statistics

Compare the three sets of means and standard deviations. Are they similar?

The data shows that the sample averages are close to the full dataset average but not exact. The standard deviations also vary, with the full dataset having the most variation, while the smaller samples have lower variation.

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

This tells me that small samples don’t always match the full dataset exactly, but they can still give a good idea. The bigger the sample, the closer it usually gets to the full dataset. This shows that random sampling works well but always has some differences.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
