
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
```


```{r}
data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQ_02vV4-U327jbHQCTXZGpnDUyW46hKPwMJwVVZXcLeefOo2JR6iR5SPs-GuiK9ZfekcUfvwbTp8eJ/pub?output=csv")
```

```{r}
data_long <- data %>%
  pivot_longer(cols = -Year, names_to = "Format", values_to = "Revenue") %>%
  mutate(
    Revenue = str_replace_all(Revenue, "[$,]", ""),  # remove dollar signs and commas
    Revenue = as.numeric(Revenue)  # convert to numeric
  )

```


```{r}
data_growth <- data_long %>%
  group_by(Format) %>%
  arrange(Year) %>%
  mutate(Percent_Growth = (Revenue - lag(Revenue)) / lag(Revenue) * 100) %>%
  drop_na()
```

```{r}
cor_test <- cor.test(data_growth$Year, data_growth$Percent_Growth)
print(cor_test)

```

```{r}
table(data_growth$Format)
table(data_growth$Type)

```
```{r}
data_growth <- data_growth %>%
  mutate(Type = case_when(
    str_detect(Format, regex("vinyl|cd|physical", ignore_case = TRUE)) ~ "Physical",
    TRUE ~ "Digital"
  ))

```
```{r}
table(data_growth$Type)

```
```{r}
data_growth <- data_growth %>%
  mutate(
    Format = factor(Format),
    Type = factor(Type)
  )

```
```{r}
model <- lm(Percent_Growth ~ Year * Type, data = data_growth)
summary(model)

```

```{r}
data_growth <- data_growth %>%
  mutate(Type = case_when(
    str_detect(Format, "Vinyl|CD|Physical") ~ "Physical",
    TRUE ~ "Digital"
  ))

```

```{r}
model <- lm(Percent_Growth ~ Year * Type, data = data_growth)
summary(model)

```
```{r}
table(data_growth$Type)

```
```{r}
unique(data_growth$Format)

```




```{r}
ggplot(data_growth, aes(x = Year, y = Percent_Growth, color = Format)) +
  geom_line() +
  labs(title = "Percentage Growth of Music Revenue Formats Over Time")
```


AI Link: https://chatgpt.com/share/6816d502-fcb8-800a-bb94-a32956efe3b9

I've gotten quite a bit of error codes, and even when I asked ChatGPT about it. I still could not get it solved in order to get correlation, ANOVA, and regression of my data. Even after fixing up my dataset.However, I was able to get information for correlation. Unfortunately, all it told me was there is no significant linear relationship between Year and Percent Growth. The correlation is very weak (−0.0416) and the p-value confirms it’s not statistically significant.

I tried for over an hour to get the other two codes to workout, but even with the help of AI I still couldn't get results.

Tests: I wasn't too sure, but I do have some ideas that could possibly work for my data.

ANOVA
I can use ANOVA to compare the percentage growth rates of different categories (Vinyl, Streaming).
This tells me if the average growth is statistically different across formats.
(Are physical formats growing faster than streaming?)

Correlation 
I can calculate the correlation between Vinyl (and/or total physical) growth rates and streaming growth rates.
If I find a negative correlation, it would support the idea that as physical grows, streaming slows a little, or at least that they don't grow together.
(Are physical and streaming moving together or separately?)

Z-Scores
By standardizing the growth rates (using Z-scores), I can identify years where physical formats had especially strong growth compared to streaming.
It helps highlight unusual/important years that push my hypothesis.

Summary - 
ANOVA:	Compare % growth between physical formats and streaming
Correlation:	See if physical growth and streaming growth are linked or not
Z-Scores:	Spot years where physical growth was unusually high



Original data:

https://www.riaa.com/wp-content/uploads/2023/03/2022-Year-End-Music-Industry-Revenue-Report.pdf
https://www.riaa.com/wp-content/uploads/2024/03/2023-Year-End-Revenue-Statistics.pdf
https://www.riaa.com/wp-content/uploads/2025/03/RIAA-2024Year-End-Revenue-Report.pdf

My own iteration of the data all together:

```{r}
RIAA_revenue <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSdHiPGcxYmcvFAPLVQZ7BvEEsjBrkA3hXCwgqjcz9w5Obu0zd-a_sV7GxfsP4sU_ucjIX6nEfOIdYx/pub?output=csv")
```
What the dataset contains:
Revenue by category from 2021 to 2024 (in millions of dollars).


Categories include:

Streaming (e.g., Paid Subscription, Ad-Supported)

Physical formats (e.g., CD, LP/EP (vinyl), Music Video, Other Physical)

Totals for streaming and physical

% Change columns show year-over-year growth or decline for each category.


Time period covered:
Four years: 2021, 2022, 2023, and 2024


Key columns for analysis:
LP/EP: vinyl record sales.

CD, Music Video, Other Physical: Useful for total physical media analysis.

Total Physical Value: compare overall physical trends.

Total Streaming Revenues: Use this to contrast physical vs. digital trends.


What’s missing or needed:
Consumer behavior data (demographics, preferences) isn’t included—it’s revenue-only.

No breakdown within LP/EP (e.g., genre, age group).

Inflation adjustment if you want to analyze “real” revenue change.


Questions:
Is vinyl driving the rise in physical media overall?

Are CDs declining while vinyl rises?

How does total physical media trend compare to streaming growth?

Does the growth of vinyl mirror any dips or flattening in streaming?

